
Reproducible Research Course Project 1
=======================================

##Analysing Activity Data

1. Code for reading in the dataset and/or processing the data  

```{r}
activity <- read.csv("activity.csv")
dim(activity)
head(activity)
summary(activity)

```

2. Histogram of the total number of steps taken each day  

```{r}
stepsPerDay <- tapply(activity$steps, activity$date, sum)
hist(stepsPerDay, breaks = 10)

```

3. Mean and median number of steps taken each day  
```{r}
summary(stepsPerDay)

```
4. Time series plot of the average number of steps taken  
```{r}
meanIntervalSteps <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
summary(meanIntervalSteps)
head(meanIntervalSteps)
str(meanIntervalSteps)
plot(names(meanIntervalSteps),meanIntervalSteps, type = "l")

```

5. The 5-minute interval that, on average, contains the maximum number of steps  
```{r}
names(meanIntervalSteps)[which.max(meanIntervalSteps)]

```

6. Code to describe and show a strategy for imputing missing data  
    * Check how many rows with missing data exist
    * Missing values will be impute according to mean value of corresponding interval
    * Creating imputed data in a new column
```{r}
sum(is.na(activity$steps))
activity$steps_imp <- ifelse(is.na(activity$steps), meanIntervalSteps[as.character(activity$interval)], activity$steps)

```

7. Histogram of the total number of steps taken each day after missing values are imputed  
```{r}
stepsPerDay <- tapply(activity$steps_imp, activity$date, sum)
hist(stepsPerDay, breaks = 10)
summary(stepsPerDay)

```
It seems that the imputation did not affect the mean (as expected),  
but it caused the median to rise and become equal to the mean.

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  
```{r}
activity$weekday <- as.numeric(format(as.Date(activity$date),"%w"))
activity$is.weekend <- ifelse(activity$weekday == 6 | activity$weekday == 0, 1, 0)
wdActivity <- subset(activity, is.weekend == 0)
wdActivityIM <- tapply(wdActivity$steps_imp, wdActivity$interval, mean)
weActivity <- subset(activity, is.weekend == 1)
weActivityIM <- tapply(weActivity$steps_imp, weActivity$interval, mean)
par(mfrow = c(2,1), mar = c(5,5,2,2))
plot(names(wdActivityIM),wdActivityIM, type = "l", main = "Weekdays", xlab = "" , ylab = "Average Steps",xlim = c(0,2500))
plot(names(weActivityIM),weActivityIM, type = "l", main = "Weekends", xlab = "Interval", ylab = "Average Steps",xlim = c(0,2500))
 
```

End of Project 1