---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
stepsPerDay <- tapply(activity$steps, activity$date, sum)
hist(stepsPerDay, breaks = 10)

```

## What is mean total number of steps taken per day?

```{r}

summary(stepsPerDay)[c('Mean','Median')]

```

## What is the average daily activity pattern?

```{r}

meanIntervalSteps <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(names(meanIntervalSteps),meanIntervalSteps, type = "l")

```

The 5-minute interval that, on average, contains the maximum number of steps  
```{r}

names(meanIntervalSteps)[which.max(meanIntervalSteps)]

```

## Imputing missing values

```{r}

sum(is.na(activity$steps))
activity$steps_imp <- ifelse(is.na(activity$steps), meanIntervalSteps[as.character(activity$interval)], activity$steps)
stepsPerDay <- tapply(activity$steps_imp, activity$date, sum)
hist(stepsPerDay, breaks = 10)
summary(stepsPerDay)

```
It seems that the imputation did not affect the mean (as expected),  
but it caused the median to rise and become equal to the mean.

## Are there differences in activity patterns between weekdays and weekends?

```{r}

activity$weekday <- as.numeric(format(as.Date(activity$date),"%w"))
activity$is.weekend <- ifelse(activity$weekday == 6 | activity$weekday == 0, 1, 0)
wdActivity <- subset(activity, is.weekend == 0)
wdActivityIM <- tapply(wdActivity$steps_imp, wdActivity$interval, mean)
weActivity <- subset(activity, is.weekend == 1)
weActivityIM <- tapply(weActivity$steps_imp, weActivity$interval, mean)
par(mfrow = c(2,1), mar = c(5,5,2,2))
plot(names(wdActivityIM),wdActivityIM, type = "l", main = "Weekdays", xlab = "" , ylab = "Average Steps",xlim = c(0,2500))
plot(names(weActivityIM),weActivityIM, type = "l", main = "Weekends", xlab = "Interval", ylab = "Average Steps",xlim = c(0,2500))
 
```
